### 2017-02-22 short-paper/clickstream-hmm-short.tex
Add L@S17 short paper.

\documentclass{sigchi-ext}

% Please be sure that you have the dependencies (i.e., additional
% LaTeX packages) to compile this example.
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage[scaled=.92]{helvet} % for proper fonts
\usepackage{graphicx} % for EPS use the graphics package instead
\usepackage{balance}  % for useful for balancing the last columns
\usepackage{booktabs} % for pretty table rules
\usepackage{ccicons}  % for Creative Commons citation icons
\usepackage{ragged2e} % for tighter hyphenation

% Some optional stuff you might like/need.
% \usepackage{marginnote}
% \usepackage[shortlabels]{enumitem}
% \usepackage{paralist}
\usepackage[utf8]{inputenc} % for a UTF8 editor only
\usepackage{amsmath}
\usepackage{float}
\usepackage[numbers]{natbib}

\usepackage{subcaption}
\usepackage{pdflscape}
\usepackage{afterpage}

\renewcommand\cite{\citep}
\def\meta/{\textsc{MeTA}}
\newcommand{\textretrieval}{textretrieval-001}
\newcommand{\sustain}{sustain-001}
\newcommand{\UIUC}{UIUC}

%% EXAMPLE BEGIN -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP --
% \copyrightinfo{Permission to make digital or hard copies of all or
% part of this work for personal or classroom use is granted without
% fee provided that copies are not made or distributed for profit or
% commercial advantage and that copies bear this notice and the full
% citation on the first page. Copyrights for components of this work
% owned by others than ACM must be honored. Abstracting with credit is
% permitted. To copy otherwise, or republish, to post on servers or to
% redistribute to lists, requires prior specific permission and/or a
% fee. Request permissions from permissions@acm.org.\\
% {\emph{CHI'14}}, April 26--May 1, 2014, Toronto, Canada. \\
% Copyright \copyright~2014 ACM ISBN/14/04...\$15.00. \\
% DOI string from ACM form confirmation}
%% EXAMPLE END

\copyrightinfo{Permission to make digital or hard copies of all or part of
  this work for personal or classroom use is granted without fee provided
  that copies are not made or distributed for profit or commercial
  advantage and that copies bear this notice and the full citation on the
  first page. Copyrights for components of this work owned by others than
  the author(s) must be honored. Abstracting with credit is permitted. To
  copy otherwise, or republish, to post on servers or to redistribute to
  lists, requires prior specific permission and/or a fee. Request
  permissions from Permissions@acm.org.\\
  \emph{L@S 2017}, April 20 - 21, 2017, Cambridge, MA, USA.\\
  Copyright is held by the owner/author(s). Publication rights licensed to
  ACM.\\
  ACM 978-1-4503-4450-0/17/04 \$15.00\\
  DOI: \url{http://dx.doi.org/10.1145/3051457.3053986}}

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{Modeling MOOC Student Behavior with Two-Layer Hidden Markov
Models}
\def\plainauthor{Chase Geigle, ChengXiang Zhai}
\def\emptyauthor{}
\def\plainkeywords{MOOC log analysis; studetn behavior modeling; Markov models; hidden Markov models}
\def\plaingeneralterms{\plainkeywords}

\title{\plaintitle}

\numberofauthors{2}
% Notice how author names are alternately typesetted to appear ordered
% in 2-column format; i.e., the first 4 autors on the first column and
% the other 4 auhors on the second column. Actually, it's up to you to
% strictly adhere to this author notation.
\author{%
  \alignauthor{%
    \textbf{Chase Geigle}\\
    \affaddr{University of Illinois at Urbana-Champaign} \\
    \affaddr{201 N Goodwin Ave}\\
    \affaddr{Urbana, IL 61801} \\
    \email{geigle1@illinois.edu} } \vfil \alignauthor{%
    \textbf{ChengXiang Zhai}\\
    \affaddr{University of Illinois at Urbana-Champaign} \\
    \affaddr{201 N Goodwin Ave}\\
    \affaddr{Urbana, IL 61801} \\
    \email{czhai@illinois.edu} } }

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},
%  pdfauthor={\plainauthor},
  pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
}

% \reversemarginpar%

\newcommand{\citemeta}{\cite{Massung:2016:ACL}}

\begin{document}

\maketitle

% Uncomment to disable hyphenation (not recommended)
% https://twitter.com/anjirokhan/status/546046683331973120
\RaggedRight{}

% Do not change the page size or page settings.
\begin{abstract}
  Massive open online courses (MOOCs) provide educators with an abundance
  of data describing how students interact with the platform, but this data
  is highly underutilized today. This is in part due to the lack of
  sophisticated tools to provide interpretable and actionable summaries of
  huge amounts of MOOC activity present in log data. In this paper, we
  propose a method for automatically discovering student behavior patterns
  by leveraging the click log data that can be obtained from the MOOC
  platform itself in a completely unsupervised manner.
\end{abstract}

\keywords{\plainkeywords}

\begin{CCSXML}
  <ccs2012>
    <concept>
      <concept_id>10002950.10003648.10003670.10003683</concept_id>
      <concept_desc>Mathematics of computing~Kalman filters and hidden Markov models</concept_desc>
      <concept_significance>500</concept_significance>
    </concept>
    <concept>
      <concept_id>10010147.10010257.10010258.10010260.10010267</concept_id>
      <concept_desc>Computing methodologies~Mixture modeling</concept_desc>
      <concept_significance>500</concept_significance>
    </concept>
  </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Mathematics of computing~Kalman filters and hidden Markov models}
\ccsdesc[500]{Computing methodologies~Mixture modeling}

\printccsdesc{}

\input{sec-intro}
\input{sec-model}
\input{sec-results}
\input{sec-conclusion}

\section{Acknowledgments}

This material is based upon work supported by the NSF GRFP under Grant
Number DGE-1144245.

\balance{}

\bibliographystyle{acm-sigchi-modern}
\bibliography{bib}

\end{document}