### 2017-02-22 short-paper/sec-conclusion.tex

\section{Conclusions and Future Directions}
We proposed a two-layer hidden Markov model for MOOC student behavior
modeling on large collections of actions extracted from MOOC log data. This
model uniquely treats behavior patterns as a sequence of \emph{latent states},
rather than assigning these states in a rule-based manner, captures the
variable behaviors of students over time, and allows analysis at different
levels of granularity.

Although we only experimented with our model on one MOOC, the model is
completely general and can be easily applied to analyze the log of any
other course to enable deep understanding of student behaviors as well as
the correlations of such behaviors as other variables such as grades.
Our model can be used to produce an ordered list of latent states for each
student over time from which we could correlate course-wide drifts with
events in a course, which could for example allow the automatic discovery
of difficult segments of a course.