### 2017-02-22 short-paper/sec-intro.tex

\section{Introduction}
The proliferation of massive open online courses (MOOCs) has resulted in a
profound impact on education. As more and more learners turn to MOOCs to
educate themselves on various topics, more and more behavioral data is
being collected as part of the system on which the MOOC is offered. The
data present in these logs has the power to aid us in understanding the
behavior of students who take our MOOCs, which is mostly undetectable for
instructors of these MOOCs today due to its vast scale. As a result, the
rich data available through these MOOC logs is highly underutilized today.

In this paper, we propose a novel two-layer hidden Markov model
(TL-HMM) to discover latent student behavior patterns via unsupervised
learning on large collections of student behavior observation sequences.
This \emph{latent variable approach} to mining student behavior
patterns that is \emph{probabilitic} for inference and \emph{flexible to
model state changes over different time resolutions} gives our model
a unique perspective compared to previous work that uses rule-based
approaches~\citep{Kizilcec:2013:LAK,Davis:2016:EDM} and work that comes
with a particular assumption about the resolution of times
considered~\citep{Faucon:2016:EDM, Kizilcec:2013:LAK}.