### 2017-02-22 short-paper/clickstream-hmm-short.tex
Update short paper to use SIGCHI Proceedings format.

%% SIGCHI Proceedings Format (modern) Sample Document
%% This sample file demonstrates the use of the `sigchi-modern' class file.
%%
%% The content of this document draws heavily on the `HCI Archive Format'
%% template supplied by SIGCHI, but has been modified to demonstrate and
%% document the LaTeX class.

% The parameter given to the `\documentclass' command can be one of:
%   `preprint', `submission', or `final'.
% These control the presence of certain features for each stage of the authoring
% process. For example, `submission' does not output the author block, and
% `final' disables page numbers.
%
% The \preprintonly{}, \submissiononly{}, and \finalonly{} commands allow you
% to wrap text that will only be output under the corresponding option.

\PassOptionsToPackage{inline}{enumitem}
\documentclass[final]{sigchi-modern}

\def\meta/{\preprintonly{\textsc{MeTA}}\finalonly{\textsc{MeTA}}%
\submissiononly{(redacted)}}

\newcommand{\citemeta}{\preprintonly{\cite{Massung:2016:ACL}}%
  \finalonly{\cite{Massung:2016:ACL}}%
  \submissiononly{[redacted citation]}}

% These packages aren't required, but do provide some helpful features that are
% used in this sample.
\usepackage{graphicx}  % Graphics for figures
\usepackage{booktabs}  % Nice formatting for tables
\usepackage{balance}   % Attempts to balance the columns on the last page

% A nice way to help avoid overful lines handing out into the margins is the
% `microtype' package, which lets certain characters extrude a *little* bit.
\usepackage{microtype}
% If the situation becomes dire, use:
%\sloppy

% You may use any bibliography package as long as the output matches that of
% the prescribed format. `natbib' is a fairly nice default, and the provided
% BibTeX style file has been tested with it.
\usepackage[numbers]{natbib}
\bibliographystyle{acm-sigchi-modern}

\usepackage{subcaption}
\usepackage{pdflscape}
\usepackage{afterpage}

% Define these tokens as appropriate for the conference to generate the
% necessary copyright notice for `final` mode.
\confname{L@S 2017}
\confdate{April 20-21}
\confyear{2017}
\conflocation{Cambridge, MA, USA}
\procissn{978-1-4503-4450-0/17/04}
\doi{10.1145/3051457.3053986}
% \copylicense be set to one of: \acmcopyright, \authorlicense, or \openlicense
%\copylicense{\authorlicense}

% End of preamble. Here comes the document.
\begin{document}

\title{Modeling MOOC Student Behavior With Two-Layer Hidden Markov Models}

% Author information can be set automatically if you provide a series of
% author names and affiliations:
%   \author[1]{Author One}
%   \author[2]{Author Two}
%   \author[1,2]{Author Three}
%   \affiliation[1]{Organisation One\\
%                   Some Country}
%   \affiliation[2]{Organisation Two\\
%                   Some Country}
%   \authorextra{\mailto{one@example.com}, \mailto{two@example.org},
%                \mailto{three@example.net}}
% Or, manually:
%   \author{
%     \authorname{Author One}\\
%     \authoraffil{
%       Organisation One\\
%       Some Country\\
%       \mailto{one@example.com}}}
%   \author{
%     \authorname{Author Two}\\
%     \authoraffil{
%       Organisation Two\\
%       Some Country\\
%       \mailto{two@example.org}}}
% Refer to the body text of the document for more details.
\author[1]{Chase Geigle}
\author[1]{ChengXiang Zhai}

% Set \authorwidth with the longest author name if you don't like the default
% spacing:
\settowidth{\authorwidth}{ChengXiang Zhai1}
\settowidth{\affilwidth}{University of Illinois at Urbana-ChampaignAAA}

\affiliation[1]{Department of Computer Science\\
    University of Illinois at Urbana-Champaign\\
    Urbana, Illinois, USA}

\authorpostscript{\email{\{%
  \href{mailto:geigle1@illinois.edu}{geigle1},
  \href{mailto:czhai@illinois.edu}{czhai}\}@illinois.edu}}

% If you format the author blocks manually, use \authorlist with a
% comma-delimited list of the author names.
% This is written into the PDF metadata for preprint and final modes.
%\authorlist{Author One, Author Two, Author Three, Author Four}

% Produces the title, author, and optional banner block.
\maketitle

\begin{abstract}
  Massive open online courses (MOOCs) provide educators with an abundance
  of data describing how students interact with the platform, but this data
  is highly underutilized today. This is in part due to the lack of
  sophisticated tools to provide interpretable and actionable summaries of
  huge amounts of MOOC activity present in log data. In this paper, we
  propose a new method for automatically discovering student behavior
  patterns by leveraging the click log data that can be obtained from the
  MOOC platform itself in a completely unsupervised manner.
\end{abstract}

% A semicolon-separated list of keywords to describe the paper.
\keywords{MOOC log analysis; student behavior modeling; Markov models;
hidden Markov models}

\newcommand{\acknowledgments}{This material is based upon work supported by
the National Science Foundation under Grant Number DGE-1144245 (GRFP) and
Grant Number 1629161.}

% ACM classification block.
% \category{}{}{}[] has three mandatory arguments and one optional argument; it
% can be used multiple times.
%\classification{%
%    \category{}{}{}
%}

\begin{CCSXML}
  <ccs2012>
    <concept>
      <concept_id>10002950.10003648.10003670.10003683</concept_id>
      <concept_desc>Mathematics of computing~Kalman filters and hidden Markov models</concept_desc>
      <concept_significance>500</concept_significance>
    </concept>
    <concept>
      <concept_id>10010147.10010257.10010258.10010260.10010267</concept_id>
      <concept_desc>Computing methodologies~Mixture modeling</concept_desc>
      <concept_significance>500</concept_significance>
    </concept>
  </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Mathematics of computing~Kalman filters and hidden Markov models}
\ccsdesc[500]{Computing methodologies~Mixture modeling}

\printccsdesc{}

\input{../sec-intro}
\input{sec-model}
\input{sec-results}
\input{sec-conclusion}

\section{Acknowledgments}
\preprintonly{\acknowledgments{}}
\finalonly{\acknowledgments{}}
\submissiononly{(Anonymized for blind reviewing.)}

\balance
\bibliography{bib}

\end{document}