### 2017-02-22 short-paper/sec-conclusion.tex

\section{Conclusion}
% summarize the main contributions of the paper; summarize the experimental
% results; argue again for our novelty; discuss future directions
We proposed a two-layer hidden Markov model for MOOC student behavior
modeling in an unsupervised learning framework on large collections of
actions extracted from MOOC log data. This model is different from existing
methods in that it treats behavior patterns as a sequence of \emph{latent
states}, rather than assigning these states in a rule-based manner. It
captures the variable behaviors of students over time, and allows
analysis at different levels of granularity.

We showed that such a model does in fact capture meaningful behavior
patterns and produces descriptions of these behavior patterns that are easy
to interpret. We argued that it is important to capture student behavior
patterns with more sophisticated models than simple discrete distributions
over actions in order to capture information present in bigrams of actions
(or above). Finally, we investigated whether we can detect differences in
student behavior patterns as they correlate with course performance.
Specifically, we demonstrated that high-performing students produce
substantially different HMM transition diagrams that tend to show longer
concentration span in quiz-taking and more active forum participation as
compared with the average students.

\subsection{Future Work}

Although we only experimented with our model on one MOOC, the model is
completely general and can be easily applied to analyze the log of any
other course to enable deep understanding of student behaviors as well as
the coorrelations of such behaviors as other variables such as grades. We
plan to develop a MOOC log analysis system based on the proposed model to
both facilitate education research and help instructors improve course
design.

Specifically, our model can be used to produce an ordered list of latent
states for each student over time. Given these labeled sequences, we could
correlate course-wide drifts in these latent states with events in a
course. For example, we might be able to automatically discover difficult
or confusing parts of a course by noticing spikes in the distribution of
students over latent states over time.

%Our model does not explicitly model drop-out like
%\citet{Kizilcec:2013:LAK}, but doing so is an obvious extension. Our model
%should be able to provide predictions of when a student is ``at risk'' for
%dropping out under such an extension.
%
%Currently, the model learns a transition matrix over the latent states that
%is \emph{shared} across all students. It would be interesting to instead
%learn a different latent state transition matrix for each individual
%student, but keep the second-level Markov models shared. This would provide
%the model with more flexibility which may be desirable itself, but would
%also naturally result in a description of a student (via his or her HMM
%transitions) that could be incorporated into existing supervised learning
%techniques that try to predict student outcomes for understanding
%which of the latent behavior patterns discovered by TL-HMM are most predictive
%of the performance of student learning.