### 2017-06-14 jedm-reviews/round-2/reviewer_b_response.md
Update paper for reviewer b; add response.

> Related works
>
> In prior iterations, the related works did not frame the authors' research
> with respect to similar methods. I think the authors have done an excellent
> job in reworking this section. It is much clearer how the authors' work is
> novel in both the education and data mining/machine learning domains.

Thank you. We greatly appreciate your helpful feedback that helped improve
this aspect of the paper's presentation.

> Abstract and style
>
> The abstract is now sufficiently long and adequately written. The table
> captions have also been fixed. Certain parts of the figures are still hard
> to read, but I understand that the possible fixes here are limited. In that
> regard, the visual issues do not significantly hinder interpretability of
> the concepts.

We again thank you for your help in improving the quality of our abstract.

> Challenges / Limitations
>
> The authors have created a full challenges and limitations section. The
> authors have added a great deal of clarity to the core issues they
> encountered with both the methods and outcomes of their work.

We agree that this added section brings much more clarity to the technical
challenges and potential limitations of our approach, and thank you for the
feedback leading to this improvement to the paper.

> General readability comments:
>
> Previously, I made comments about the readability of the work and the
> authors requested a few examples. I have written down a few quotes from the
> paper along with explanations or fixes for the problems (denoted by an
> arrow ->). It is important to note that these are not all of the issues in
> the paper. I think that authors could still make many improvements to
> increase both the readability and clarity of the work.

Thank you for the detailed list of suggestions. In the revision, we have
incorporated these suggested changes to the wording and have also done
another editing pass to attempt to correct additional errors of similar
form. We hope that the readability and clarity of the work has improved as
a direct result.