### 2017-08-28 sec-conclusion.tex
Address copy-editing feedback.

\section{Conclusions and Future Work}
As a tool to help instructors and education researchers better understand
the behavior of MOOC students, we proposed a two-layer hidden Markov model
to automatically extract student activity patterns in the form of behavior
state-transition graphs from large amounts of MOOC log data. This model is
different from existing methods in that it treats behavior patterns as a
sequence of \emph{latent states}, rather than assigning these states in a
rule-based manner. It captures the variable behaviors of students over time
and allows analysis at different levels of granularity.

We showed that such a model does, in fact, capture meaningful behavior
patterns and produces descriptions of these behavior patterns that are easy
to interpret. We argued that it is important to capture student behavior
patterns with more sophisticated models than simple discrete distributions
over actions to capture information present in bigrams of actions (or
larger sequences). By varying the number of latent states inferred, we
showed that the model is flexible and can capture patterns of differing
levels of specificity in this way.  Finally, we investigated whether we can
detect differences in student behavior patterns as they correlate with
course performance. Specifically, we demonstrated that high-performing
students produce substantially different HMM transition diagrams that tend
to show longer concentration span in quiz-taking and more active  forum
participation as compared with the average students.  These results show
the great potential of the proposed model for serving as a tool to help
humans discover knowledge about student behaviors.

%\subsection{Future Work}

There are a number of interesting directions to further extend our work in
the future.  First, the proposed model is completely general and can thus
be easily applied to analyze the log of many other courses to enable deep
understanding of student behaviors as well as the correlations of such
behaviors and other variables such as grades. To realize these benefits, it
would be useful to develop a MOOC log analysis system based on the proposed
model to facilitate education research and help instructors improve course
design.

Second, our model can empower many new comparative analyses. For example,
we could now look at how behavior patterns change between different
offerings of the same MOOC to understand how changes in course structure or
materials influence student behavior. Individual students can now also be
compared against each other or against groups. For example, by decoding the
latent state sequences for each student, we can measure how ``surprising''
their latent state transition sequence is relative to the average we would
expect according to the model, or to the average ``perfect'' student, etc.
We can now investigate how certain behavioral patterns correlate with
properties of a student (e.g., demographics, prior aptitude, etc.). After
decoding the students' latent state sequences, we could also correlate
course-wide drifts in these latent states with events in a course. For
example, we might be able to automatically discover difficult or confusing
parts of a course by noticing spikes in the distribution of students over
latent states over time.

Third, there is more recent work on better learning algorithms for mixtures of
Markov models~\citep{Gupta:2016:NIPS}. It would be worth exploring whether
the advances proposed in this and similar work can be applied to our model
to address some of the concerns surrounding our use of the EM algorithm for
our parameter estimation.

Finally, the proposed model can be extended in several ways. For example,
although our model does not explicitly model drop-out like
\citet{Kizilcec:2013:LAK}, doing so is an obvious extension. Our model
would be able to provide predictions of when a student is ``at risk'' for
dropping out under such an extension. Also, currently, the model learns a
transition matrix over the latent states that is \emph{shared} across all
students. It would be interesting to instead learn a different latent state
transition matrix for each individual student, but keep the second-level
Markov models shared. This would provide the model with more flexibility
which may be desirable itself, but would also naturally result in a
description of a student (via his or her HMM transitions) that could be
incorporated into existing supervised learning techniques that try to
predict student outcomes for understanding which of the latent behavior
patterns discovered by 2L-HMM are most predictive of the performance of
student learning. One could also relax this somewhat and posit that
\emph{groups} of students transition between the lower layer patterns
identified by our method in distinct ways; this way, we can do a soft
clustering of students into $K_2$ groups based on the similarity of their
transitioning behavior between the higher-level behaviors we can identify.